---
title: "Make Phase Eval Target Beds"
author: "Nate Olson"
date: '`r Sys.Date()`'
output: 
    bookdown::html_document2:
        toc: true
        toc_float: true
        df_print: paged
        code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
```


# Objective
Generate bed files with easy regions for use in validating phasing results. 


Targeted regions, intersection of refseq CDS and V3.3.2 bed
- Input bed: Intersection of refseq CDS, and V3.3.2 bed (not including V0.6 bed as there is no comparable bed for 38)

# Inputs
`phase_eval_targets_*bed` files generated using `make_phase_eval_bed.sh`. 

# Subset target region selection. 

## GRCh37
```{r}
phase_target_37 <- here("resources","phase_eval_targets_HG002_GRCh37.bed") %>% 
    read_tsv(col_types = "cdd", col_names = c("CHROM", "START", "END"))

## Selecting 10 regions per chrom
set.seed("53117")
target_subset <- phase_target_37 %>% 
    mutate(SIZE = END - START) %>% 
    group_by(CHROM) %>% 
    top_n(10, SIZE)

subset_targets_bed <- here("resources","phase_eval_targets_subset_GRCh37.bed")
target_subset %>% select(-SIZE) %>% 
    write_tsv(subset_targets_bed, 
              col_names = FALSE)

## Writing to vcf for manual curation
write_csv(target_subset, here("data","phase_manual_curation_GRCh37.csv"))
```

bgzip and tabix for IGV
```{bash}
bgzip -f ../resources/phase_eval_targets_subset_GRCh37.bed
tabix ../resources/phase_eval_targets_subset_GRCh37.bed.gz
```

Creating batch script to generate IGV snapshots
```{bash}
bedtools igv \
    -i ../resources/phase_eval_targets_subset_GRCh37.bed.gz \
    -path ../data/phase_manual_curation/GRCh37 \
    -sess ../resources/igv_session_giab-ul-GRCh37.xml \
    -slop 1000 \
    > ../scripts/phase_eval_GRCh37_igv.batch
```

## GRCh38
```{r}
phase_target_38 <- here("resources","phase_eval_targets_HG002_GRCh38.bed") %>% 
    read_tsv(col_types = "cdd", col_names = c("CHROM", "START", "END"))

## Selecting 10 regions per chrom
set.seed("53117")
target_subset <- phase_target_38 %>% 
    mutate(SIZE = END - START) %>% 
    mutate(CHROM = factor(CHROM, levels = paste0("chr",1:22))) %>% 
    group_by(CHROM) %>% 
    top_n(10, SIZE) 


subset_targets_bed <- here("resources","phase_eval_targets_subset_GRCh38.bed")
target_subset %>% select(-SIZE) %>% 
    write_tsv(subset_targets_bed, 
              col_names = FALSE)

## Writing to vcf for manual curation
write_csv(target_subset, here("data","phase_manual_curation_GRCh38.csv"))
```

bgzip and tabix for IGV
```{bash}
bgzip -f ../resources/phase_eval_targets_subset_GRCh38.bed
tabix ../resources/phase_eval_targets_subset_GRCh38.bed.gz
```

Creating batch script to generate IGV snapshots
```{bash}
bedtools igv \
    -i ../resources/phase_eval_targets_subset_GRCh38.bed.gz \
    -path ../data/phase_manual_curation/GRCh38 \
    -sess resources/igv_session_giab-ul-GRCh38.xml \
    -slop 1000 \
    > ../scripts/phase_eval_GRCh38_igv.batch
```


# System Information

```{r}
s_info <- devtools::session_info()
print(s_info$platform)
```
```{r}
as.data.frame(s_info$packages) %>% 
  filter(attached) %>% 
  select(package, loadedversion, date, source) %>% 
  knitr::kable()
```

## Software Versions
__BEDtools__
```{r}
system("bedtools --version",intern = TRUE)
```